---
output:
  pdf_document:
    toc: TRUE
---

```{r setup, include=FALSE}

require(rmarkdown)
require(knitr)
require(tinytex)
require(kableExtra)
require(float)
require(RColorBrewer)
#install_tinytex()

require(tidyverse)
require(stringr)
require(treemap)
require(maditr)

options(knitr.table.format = "latex")
knitr::opts_knit$set(root.dir = path.to.scripts)
knitr::opts_chunk$set(echo = FALSE)

raab<-read.csv(paste0(path.to.raab,"raabs_618.csv"))
raab<-raab[raab$raab_id==ID,]

pop<-read.csv(paste0(path.to.pop,"raabs_pop_618_peek_format.csv"))
popfives<-pop[pop$raab_id==ID,]

meta<-read.csv(paste0(path.to.raab,"raab-log_v3.csv"))
raab_meta<-meta[meta$raab_id==ID,]
PI<-raab_meta$pi_name
TITLE<-raab_meta$survey_title

write.table(raab_meta,file=paste0(outdir,"/raw/meta.csv"),row.names=F,col.names=T,sep=",")
write.table(raab,file=paste0(outdir,"/raw/data/survey_data.csv"),row.names=F,col.names=T,sep=",")
write.table(popfives,file=paste0(outdir,"/raw/data/population.csv"),row.names=F,col.names=T,sep=",")

source("raab_functions_v210719.R")

```

---
title: Rapid Assessment of Avoidable Blindness
subtitle: Report of findings from `r TITLE`
author: Principal Investigator `r PI`
date: Report generated `r format(Sys.time(), '%d %B %Y')`
output:
  pdf_document: 
    toc: TRUE
header-includes:
- \usepackage{float}
- \usepackage{tabu}
tables:
  yes
---

```{r raab_report_setup, include=FALSE}

source("RAAB5_report_setup.R")

```

\newpage

# Sample representation

## Table 1

```{r sum1, echo=FALSE}

source("SUM_1.R")

sum1.nice.names<-sum1

write.table(sum1.nice.names,file=paste0(outdir,"/summary/data/t1.csv"),row.names=F,col.names=T,sep=",")

names(sum1.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(sum1.nice.names, booktabs=T, linesep = "", caption = "Eligible persons, coverage, absentees and refusals") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Exam status" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("The response rate is the percent examined",notation="symbol")

```

The response rate indicates the proportion of the eligible sample who were examined.  The sample size calculator includes the expected non-response rate in the sample and increases the sample size accordingly. This ensures that the sample size is powerful enough to estimate the proposed condition with the expected accuracy. If the response rate is lower than 80-90%, there is a concern that the condition under review in the 10-20% who were not examined may be different to those that were examined. If the response rate is close to 100%, this might be an indication that eligible participants who were absent or refused were replaced by eligible participants in the next household, which would introduce bias and mean that results are less representative. The reason is that people with impaired vision are more likely to stay at home and people with good vision are more likely to be away. Absent participants should not be replaced with participants who are at home, because that would increased the chance of including more people with vision problems and falsely inflate the outcome.

Conversely, in certain settings participants may be more compliant with leadership requests to stay home, or more likely to work in the environment surrounding their home. It is important to review this information in relation to the tables on representativeness of the sample below to identify whether a high response rate is valid.

## Table 2, table 3

```{r asa1, echo=FALSE}

source("ASA_1.R")

asa1.nice.names<-asa1[,c("age.groups.tens","female.examined.n","female.examined.pct","male.examined.n","male.examined.pct","total.examined.n","total.examined.pct")]

write.table(asa1.nice.names,file=paste0(outdir,"/summary/data/t2.csv"),row.names=F,col.names=T,sep=",")

names(asa1.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(asa1.nice.names, booktabs=T, linesep = "", caption = "Age and sex distribution of people examined in the sample") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Age group (years)" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)


```


```{r asa2, echo=FALSE}

source("ASA_2.R")

asa2.nice.names<-asa2

write.table(asa2.nice.names,file=paste0(outdir,"/summary/data/t3.csv"),row.names=F,col.names=T,sep=",")

names(asa2.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(asa2.nice.names, booktabs=T, linesep = "", caption = "Total number of people aged 50+ in survey area") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Exam status" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

## Plot 1, plot 2

```{r asa1_2_plots, echo=FALSE, fig.height=3.25, fig.fullwidth=T}

fasa12<-rbind(as.numeric(asa1$female.examined.pct[1:4]),as.numeric(asa2$female.pct[1:4]))
masa12<-rbind(as.numeric(asa1$male.examined.pct[1:4]),as.numeric(asa2$male.pct[1:4]))

par(mfrow=c(1,2))

bp1<-barplot(as.matrix(fasa12),beside=T,horiz=T,yaxt='n',xaxt='n',col=c("grey","purple"),border = "white",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,16))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=c(2,5,8,11,30),labels = NA,tick=F)
text(x=par("usr")[3]-6, y=c(2,5,8,11), labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col=c("purple","grey"),pch=15,c("Survey population","Census"),bty='n',cex=0.6,title = expression(bold("Survey females")),horiz=T,title.adj=0)

bp2<-barplot(as.matrix(masa12),beside=T,horiz=T,yaxt='n',xaxt='n',col=c("grey","gold"),border = "white",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,16))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=c(2,5,8,11,30),labels = NA,tick=F)
text(x=par("usr")[3]-6, y=c(2,5,8,11), labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col=c("gold","grey"),pch=15,c("Survey population","Census"),bty='n',cex=0.6,title=expression(bold("Survey males")),horiz=T,title.adj=0)

```

```{r asa1_extra_plot, echo=FALSE, fig.height=3.25, fig.fullwidth=T}

fasa1x<-as.matrix(t(asa1[1:4,c("female.examined.pct","female.nonresponse.pct")]))
masa1x<-as.matrix(t(asa1[1:4,c("male.examined.pct","male.nonresponse.pct")]))

par(mfrow=c(1,2))

bp3<-barplot(fasa1x,horiz=T,yaxt='n',xaxt='n',col=c("white","purple"),border = "purple",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,6))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=bp3[1:4],labels = NA,tick=F)
text(x=par("usr")[3]-6, y=bp3[1:4], labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col="purple",pch=c(0,15),c("Examined","Nonresponders"),bty='n',cex=0.6,title = expression(bold("Enumerated females")),horiz=T,title.adj=0)

bp4<-barplot(masa1x,horiz=T,yaxt='n',xaxt='n',col=c("white","gold"),border = "gold",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,6))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=bp4[1:4],labels = NA,tick=F)
text(x=par("usr")[3]-6, y=bp4[1:4], labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col="gold",pch=c(0,15),c("Examined","Nonresponders"),bty='n',cex=0.6,title=expression(bold("Enumerated males")),horiz=T,title.adj=0)

```

For your results to be useful for planning, your sample needs to be representative of your population. By comparing the age-sex composition of your sample, to the age-sex composition of the population, we can assess whether the sample is representative. You can also use the age-sex composition of the population to weight your crude estimates, providing age-sex adjusted (ASA) estimates that are more representative than the crude sample outputs.

Important: if your sample differs from the population because one group was more likely to be unavailable (e.g. young men, or other seasonal labourers) then you would expect to see this reflected in the response rate (i.e. more people are unavailable).  If the difference between the sample and the population is high, but the proportion of people unavailable is low (e.g. response rate is still above 95%) this might be an indication that eligible participants who were absent or refused were replaced by others, which may introduce bias as explained above.

\newpage

# Prevalence and causes of vision impairment

We report the prevalence of vision impairment (VI) using presenting visual acuity (PVA) in the better eye. Presenting VA is measured with correction, if available.

## Table 4, table 5

```{r sum3, echo=FALSE}

source("SUM_3_SUM_4_RAAB5.R")

sum3.nice.names<-sum3

sum3.nice.names$vi.level<-as.character(sum3.nice.names$vi.level)
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="blind"]<-"Blind"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="severe.vi"]<-"Severe"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="moderate.vi"]<-"Moderate"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="mild.vi"]<-"Mild"

sum3.nice.names$female.adj.ci<-paste0(sum3.nice.names$female.adj.pct.lci," - ",sum3.nice.names$female.adj.pct.uci)
sum3.nice.names$male.adj.ci<-paste0(sum3.nice.names$male.adj.pct.lci," - ",sum3.nice.names$male.adj.pct.uci)
sum3.nice.names$total.adj.ci<-paste0(sum3.nice.names$total.adj.pct.lci," - ",sum3.nice.names$total.adj.pct.uci)
sum3.nice.names$female.ci<-paste0(sum3.nice.names$female.pct.lci," - ",sum3.nice.names$female.pct.uci)
sum3.nice.names$male.ci<-paste0(sum3.nice.names$male.pct.lci," - ",sum3.nice.names$male.pct.uci)
sum3.nice.names$total.ci<-paste0(sum3.nice.names$total.pct.lci," - ",sum3.nice.names$total.pct.uci)

sum3a.nice.names<-sum3.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(sum3a.nice.names,file=paste0(outdir,"/summary/data/t4.csv"),row.names=F,col.names=T,sep=",")

names(sum3a.nice.names)<-c("","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(sum3a.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of blindness (PVA <3/60), severe (PVA <6/60) and moderate (PVA <6/18) vision impairment") %>% 
  kable_styling(full_width=T,latex_options="HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)


```


```{r sum4, echo=FALSE}

sum3b.nice.names<-sum3.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(sum3b.nice.names,file=paste0(outdir,"/summary/data/t5.csv"),row.names=F,col.names=T,sep=",")

names(sum3b.nice.names)<-c("","%","95% CI","Extraplotated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magintude")

knitr::kable(sum3b.nice.names, booktabs=T, linesep = "", caption = "Adjusted prevalence and extrapolated magnitude of blindness (PVA <3/60), severe (PVA <6/60) and moderate (PVA <6/18) vision impairment") %>% 
  kable_styling(latex_options=c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)%>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```

These tables show the crude and adjusted prevalence of vision impairment by impairment level and gender. Table 5 also shows the estimated magnitude of vision impairment in the study area by gender, by multiplying the sample prevalence by the total population according to the uploaded population data. The 95% confidence interval is calculated to account for cluster sampling.

## Table 6, table 7

```{r asa6, echo=FALSE}

source("ASA_6_RAAB5.R")

asa6.nice.names<-asa6
asa6.nice.names$vi.level<-as.character(asa6.nice.names$vi.level)

asa6.nice.names$vi.level[asa6.nice.names$vi.level=="mild.cumulative"]<-"Mild or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="moderate.cumulative"]<-"Moderate or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="severe.cumulative"]<-"Severe or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="blind.cumulative"]<-"Blind"

asa6.nice.names$female.adj.ci<-paste0(asa6.nice.names$female.adj.pct.lci," - ",asa6.nice.names$female.adj.pct.uci)
asa6.nice.names$male.adj.ci<-paste0(asa6.nice.names$male.adj.pct.lci," - ",asa6.nice.names$male.adj.pct.uci)
asa6.nice.names$total.adj.ci<-paste0(asa6.nice.names$total.adj.pct.lci," - ",asa6.nice.names$total.adj.pct.uci)
asa6.nice.names$female.ci<-paste0(asa6.nice.names$female.pct.lci," - ",asa6.nice.names$female.pct.uci)
asa6.nice.names$male.ci<-paste0(asa6.nice.names$male.pct.lci," - ",asa6.nice.names$male.pct.uci)
asa6.nice.names$total.ci<-paste0(asa6.nice.names$total.pct.lci," - ",asa6.nice.names$total.pct.uci)


asa6a.nice.names<-asa6.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(asa6a.nice.names,file=paste0(outdir,"/summary/data/t6.csv"),row.names=F,col.names=T,sep=",")

names(asa6a.nice.names)<-c("","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(asa6a.nice.names, booktabs=T, linesep = "", caption = "Crude cumulative prevalence of blindness, severe, moderate and mild vision impairment") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold = T) %>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```



```{r asa6b,echo=FALSE}

asa6b.nice.names<-asa6.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(asa6b.nice.names,file=paste0(outdir,"/summary/data/t7.csv"),row.names=F,col.names=T,sep=",")

names(asa6b.nice.names)<-c("","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude")

knitr::kable(asa6b.nice.names, booktabs=T, linesep = "", caption = "Adjusted cumulative prevalence and extrapolated magnitude of blindness, severe vision impairment and moderate vision impairment") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold = T)%>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```


## Table 8, table 9

```{r sum6, echo=FALSE}

source("SUM_6_RAAB5.R")

sum6.nice.names<-sum6[,1:9]

sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"1. Uncorrected refractive error"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"2. Uncorrected aphakia"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_cataract_untreated"]<-"3. Untreated cataract"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"4. Cataract surgical complications"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"5. Trachomatous corneal opacity"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"6. Other corneal opacity"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_phthisis" ]<-"7. Phthisis"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_onchocerciasis"]<-"8. Onchocerciasis"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_glaucoma" ]<-"9. Glaucoma"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"10. Diabetic retinopathy"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"11. Age-related macular degeneration"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_posterior_segment_disease" ]<-"12. Other posterior segment disease"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"13. Other globe or CNS abnomalities"

write.table(sum6.nice.names,file=paste0(outdir,"/summary/data/t8.csv"),row.names=F,col.names=T,sep=",")

sum6.for.tree<-sum6.numeric

names(sum6.nice.names) <- c(' ', 'n', '%', 'n', '%','n', '%','n', '%')

knitr::kable(sum6.nice.names, booktabs=T, linesep = "", caption = " Principal cause of blindness (PVA <3/60), severe (PVA <6/60) and moderate (PVA <6/18) vision impairment") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(13,hline_after=T) %>%
  column_spec(1,width='6cm') %>%
  add_header_above(c("Principal cause" = 1, "Blind" = 2, "Severe" = 2, "Moderate" = 2, "Mild" = 2),bold = T) %>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```


```{r sum6_agg, echo=FALSE}

sum6intag<-aggregate(sum6.numeric[,2:9],by=list(sum6.numeric$cause_group_1), FUN=sum,na.rm=T)
sum6intag[nrow(sum6intag)+1,2:9]<-colSums(sum6intag[1:3,2:9])
sum6intag$Group.1[4]<-"D. Avoidable (A + B + C)"
s6tag<-aggregate(sum6.numeric[,2:9],by=list(sum6.numeric$cause_group_2), FUN=sum,na.rm=T)

sum6ag<-rbind(sum6intag,s6tag)

sum6ag[,c(8,9)]<-"*"

write.table(sum6ag,file=paste0(outdir,"/summary/data/t9.csv"),row.names=F,col.names=T,sep=",")

names(sum6ag)<-c("","n", "%", "n", "%", "n","%","n","%")

knitr::kable(sum6ag, booktabs=T, linesep = c('', '', '', '\\addlinespace'), caption = " Principal cause of blindness (PVA <3/60), severe (PVA <6/60) and moderate (PVA <6/18) vision impairment, by intervention category") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Category" = 1, "Blind" = 2, "Severe" = 2, "Moderate" = 2, "Mild" = 2),bold = T)%>%
  add_footnote(c("Participants not tested at 6/12","PHC: Primary Health Care; PEC: Primary Eye Care"),notation="symbol")

```



Table 8 compares the main cause of blindness, severe vision impairment and moderate vision impairment in the person. Table 9 shows what proportion of vision impairment is attributable to treatable, preventable and posterior segment disease. From these tables the priorities for intervention can be determined. The distribution of cases of blindness in the sample are visualised below.

## Plot 3

```{r treemap_1, echop=FALSE, fig.fullwidth=T}

sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"Uncorrected refractive error"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"Uncorrected aphakia"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_cataract_untreated"]<-"Untreated cataract"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"Cataract surgical complications"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"Trachomatous corneal opacity"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"Other corneal opacity"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_phthisis" ]<-"Phthisis"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_onchocerciasis"]<-"Onchocerciasis"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_glaucoma" ]<-"Glaucoma"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"Diabetic retinopathy"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"Age-related macular degeneration"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_posterior_segment_disease" ]<-"Other posterior segment disease"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"Other globe or CNS abnomalities"

treemap(sum6.for.tree[1:13,],index="principal.cause",vSize="blind.pct",type="index",palette = "RdYlBu",title="Principal causes of bilateral blindness",position.legend="right",fontsize.title=12)

```



## Table 10, table 11

```{r prev4, echo=FALSE}

source("PREV_4.R")

prev4.nice.names<-prev4[,1:7]

prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"1. Uncorrected refractive error"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"2. Uncorrected aphakia"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_cataract_untreated"]<-"3. Untreated cataract"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"4. Cataract surgical complications"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"5. Trachomatous corneal opacity"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"6. Other corneal opacity"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_phthisis" ]<-"7. Phthisis"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_onchocerciasis"]<-"8. Onchocerciasis"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_glaucoma" ]<-"9. Glaucoma"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"10. Diabetic retinopathy"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"11. Age-related macular degeneration"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_posterior_segment_disease" ]<-"12. Other posterior segment disease"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"13. Other globe or CNS abnomalities"

write.table(prev4.nice.names,file=paste0(outdir,"/summary/data/t10.csv"),row.names=F,col.names=T,sep=",")

names(prev4.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(prev4.nice.names, booktabs=T, linesep = "", caption = "Principal cause of blindness (PVA<3/60) in males and females") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(13,hline_after=T) %>%
  column_spec(1,width='6cm') %>%
  add_header_above(c("Principal cause" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold = T)

```

This table shows the principal cause of blindness disaggregated by sex.

```{r prev4_agg, echo=FALSE}

prev4intag<-aggregate(prev4[,2:7],by=list(prev4$cause_group_1), FUN=sum,na.rm=T)
prev4intag[nrow(prev4intag)+1,2:7]<-colSums(prev4intag[1:3,2:7])
prev4intag$Group.1[4]<-"D. Avoidable (A + B + C)"
p4tag<-aggregate(prev4[,2:7],by=list(prev4$cause_group_2), FUN=sum,na.rm=T)

prev4ag<-rbind(prev4intag,p4tag)

write.table(prev4ag,file=paste0(outdir,"/summary/data/t11.csv"),row.names=F,col.names=T,sep=",")

names(prev4ag)<-c("","n", "%", "n", "%", "n","%")

knitr::kable(prev4ag, booktabs=T, linesep = c('', '', '', '\\addlinespace'), caption = " Principal cause of blindness (PVA <3/60) in bilaterally affected individuals, by gender and intervention category") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Category" = 1, "Male" = 2, "Female" = 2, "Total" = 2),bold = T)

```


## Table 12, table 13

```{r newtab1a, echo=FALSE}

source("NEW_TAB_1_RAAB5.R")

unt.nice.names<-unt

unt.nice.names$male.ci<-paste0(unt.nice.names$male.pct.lci," - ",unt.nice.names$male.pct.uci)
unt.nice.names$female.ci<-paste0(unt.nice.names$female.pct.lci," - ",unt.nice.names$female.pct.uci)
unt.nice.names$total.ci<-paste0(unt.nice.names$total.pct.lci," - ",unt.nice.names$total.pct.uci)

row.order<-c("blind.unilat","severe.unilat","moderate.unilat","mild.unilat")
unt.nice.names<-unt.nice.names %>% arrange(match(unt.nice.names$vi.level,row.order))

unt.nice.names$vi.level<-as.character(unt.nice.names$vi.level)
unt.nice.names$vi.level[unt.nice.names$vi.level=="mild.unilat"]<-"Mild in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="moderate.unilat"]<-"Moderate in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="severe.unilat"]<-"Severe in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="blind.unilat"]<-"Blind in one eye only"

unta.nice.names<-unt.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(unta.nice.names,file=paste0(outdir,"/summary/data/t12.csv"),row.names=F,col.names=T,sep=",")

names(unta.nice.names) <- c(' ', 'n', '%', '95% CI', 'n', '%', '95% CI', 'n', '%', '95% CI')

knitr::kable(unta.nice.names,booktabs=T, linesep = "",caption="Crude prevalence of blind, severe and moderate unilateral vision impairment") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("VI level" = 1, "Male" = 3, "Female" = 3, "Total" = 3),bold=T) %>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```



```{r newtab1b, echo=FALSE}


unt.nice.names$male.adj.ci<-paste0(unt.nice.names$male.adj.pct.lci," - ",unt.nice.names$male.adj.pct.uci)
unt.nice.names$female.adj.ci<-paste0(unt.nice.names$female.adj.pct.lci," - ",unt.nice.names$female.adj.pct.uci)
unt.nice.names$total.adj.ci<-paste0(unt.nice.names$total.adj.pct.lci," - ",unt.nice.names$total.adj.pct.uci)


untb.nice.names<-unt.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(untb.nice.names,file=paste0(outdir,"/summary/data/t13.csv"),row.names=F,col.names=T,sep=",")

names(untb.nice.names) <- c("", "%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude")

knitr::kable(untb.nice.names, booktabs=T, linesep = "", caption="Adjusted prevalence and extrapolated magnitude of blind, severe and moderate unilateral vision impairment") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```

Cases of unilateral vision impairment acknowledge an additional sight loss burden in the population not captured in the standard WHO definition of vision impairment which refers to bilateral impairment only.

\newpage

# Cataract

## Table 14, table 15 

```{r asa7a, echo=FALSE}

source("PREV_14_RAAB5.R")
source("ASA_7_RAAB5.R")

asa7a.int<-asa7[,c("vi.level","female.n","female.pct","female.pct.lci","female.pct.uci","male.n","male.pct","male.pct.lci","male.pct.uci","total.n","total.pct","total.pct.lci","total.pct.uci")]

asa7a.int$female.ci<-paste0(asa7a.int$female.pct.lci," - ", asa7a.int$female.pct.uci)
asa7a.int$male.ci<-paste0(asa7a.int$male.pct.lci," - ", asa7a.int$male.pct.uci)
asa7a.int$total.ci<-paste0(asa7a.int$total.pct.lci," - ", asa7a.int$total.pct.uci)

asa7a.nice.names<-asa7a.int[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(asa7a.nice.names,file=paste0(outdir,"/summary/data/t14.csv"),row.names=F,col.names=T,sep=",")

names(asa7a.nice.names)<-c(" ","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(asa7a.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of cataract with PinVA <3/60, <6/60 and <6/18") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  pack_rows("Bilateral",1,4) %>%
  pack_rows("Unilateral",5,8) %>%
  add_header_above(c("Operable cataract" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) %>%
  add_footnote(c("Participants not tested at 6/12","Unilateral refers to cases where the other eye has PVA of 6/18"),notation="symbol")


```


```{r asa7b, echo=FALSE}

source("PREV_14_RAAB5.R")
source("ASA_7_RAAB5.R")

asa7b.int<-asa7[,c("vi.level","female.adj.pct","female.adj.pct.lci","female.adj.pct.uci","male.adj.pct","male.adj.pct.lci","male.adj.pct.uci","total.adj.pct","total.adj.pct.lci","total.adj.pct.uci","extrapolated.female.n","extrapolated.male.n","extrapolated.total.n")]

asa7b.int$male.adj.ci<-paste0(asa7b.int$male.adj.pct.lci," - ",asa7b.int$male.adj.pct.uci)
asa7b.int$female.adj.ci<-paste0(asa7b.int$female.adj.pct.lci," - ",asa7b.int$female.adj.pct.uci)
asa7b.int$total.adj.ci<-paste0(asa7b.int$total.adj.pct.lci," - ",asa7b.int$total.adj.pct.uci)

asa7b.nice.names<-asa7b.int[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(asa7b.nice.names,file=paste0(outdir,"/summary/data/t15.csv"),row.names=F,col.names=T,sep=",")

names(asa7b.nice.names)<-c(" ","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude")

knitr::kable(asa7b.nice.names, booktabs=T, linesep = "", caption = "Adjusted prevalence and extrapolated magnitude of cataract with PinVA <3/60, <6/60, <6/18 and <6/12") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  pack_rows("Bilateral",1,4) %>%
  pack_rows("Unilateral",5,8) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("Operable cataract" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) %>%
  add_footnote(c("Participants not tested at 6/12","Unilateral refers to cases where the other eye has normal PVA"),notation="symbol")

```

The pinhole visual acuity (PinVA) thresholds used in Tables 14 & 15 are aligned with WHO PVA thresholds for vision impairment but refer to PinVA outcomes. In this way, they describe the prevalence and magnitude of operable cataract used in cataract surgical coverage and effective cataract surgical coverage (Table 16).

## Table 16

```{r prev14, echo=FALSE}

source("PREV_14_RAAB5.R")

prev14.nice.names<-prev14

prev14.nice.names$male.adjusted.ci<-paste0(prev14.nice.names$male.adjusted.lci," - ",prev14.nice.names$male.adjusted.uci)
prev14.nice.names$female.adjusted.ci<-paste0(prev14.nice.names$female.adjusted.lci," - ",prev14.nice.names$female.adjusted.uci)
prev14.nice.names$total.adjusted.ci<-paste0(prev14.nice.names$total.adjusted.lci," - ",prev14.nice.names$total.adjusted.uci)

prev14.nice.names$denom.thresh<-recode_factor(prev14.nice.names$denom.thresh,'612' = "<6/12",'618' = "<6/18", '660' = "<6/60",'360' = "<3/60")
#prev14.nice.names$num.thresh<-recode_factor(prev14.nice.names$num.thresh,'csc' = "CSC",'ecsc_612' = "eCSC (outcome <6/12)", 'ecsc_618' = "eCSC (outcome <6/18)")
prev14.nice.names$num.thresh<-recode_factor(prev14.nice.names$num.thresh,'csc' = "CSC")

prev14a.nice.names<-prev14.nice.names[,c("num.thresh","female.adjusted","female.adjusted.ci","male.adjusted","male.adjusted.ci","total.adjusted","total.adjusted.ci")]

write.table(prev14a.nice.names,file=paste0(outdir,"/summary/data/t16.csv"),row.names=F,col.names=T,sep=",")

names(prev14a.nice.names)<-c("Post-operative PVA","Adj. %","95% CI","Adj. %","95% CI","Adj. %","95% CI")

knitr::kable(prev14a.nice.names, booktabs=T, linesep = "", caption = "Adjusted Cataract Surgical Coverage at the person level") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(2,width='4cm') %>%
  pack_rows("Operable threshold <6/18",1,1) %>%
  pack_rows("Operable threshold <6/60",2,2) %>%
  pack_rows("Operable threshold <3/60",3,3) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("CSC: Cataract Surgical Coverage",notation="symbol")


```

CSC measures the accessibility of cataract surgical services. It describes the number of people in a population with operated cataract as a proportion of those having operable plus operated cataract. This is reported at three levels of cataract vision impairment in the denominator. Effective CSC (eCSC) is a quality-corrected version of the same metric. It describes the number of people with operated cataract and a good vision outcome of surgery as a proportion of those having operable plus operated cataract. This indicator incorporates a measure of quality along with access as an additional element of UHC. 

## Table 17

```{r sum11, echo=FALSE}

source("SUM_11.R")

sum11.nice.names<-surgery.bars.cat

write.table(sum11.nice.names,file=paste0(outdir,"/summary/data/t17.csv"),row.names=F,col.names=T,sep=",")

names(sum11.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(sum11.nice.names,booktabs=T, linesep = "", caption="Barriers to cataract surgery among participants with bilateral cataract and PinVA <6/60") %>% 
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(nrow(sum11.nice.names)-1,hline_after=T) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("Barrier" = 1, "Female" = 2, "Male" = 2, "Total" = 2)) %>%
  add_footnote("Participants can report 1 or 2 barriers each", notation="symbol")  

```

The standard RAAB survey protocol does not allow for in-depth interviews to determine why people with operable cataract have not yet been operated. This preliminary data on barriers to surgery should be regarded as an indication whether more detailed qualitative studies are required.

# Surgical outcomes

Surgical outcomes are reported for all operated eyes in the sample, not at the person level.

## Table 18

```{r newtab2, echo=FALSE}

source("OUTC_1_OUTC_2_SUM_12_RAAB5.R")

newtab2.nice.names<-newtab2

newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_iol"]<-"IOL"
newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_non_iol"]<-"Non-IOL"
newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_couching"]<-"Couching"

write.table(newtab2.nice.names,file=paste0(outdir,"/summary/data/t18.csv"),row.names=F,col.names=T,sep=",")

names(newtab2.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab2.nice.names, booktabs=T, linesep = "", caption = "Type of cataract surgery performed, count by eyes") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Surgery type" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

## Table 19, table 20

```{r newtab3a, echo=FALSE}

source("OUTC_1_OUTC_2_SUM_12_RAAB5.R")

newtab3a.nice.names<-newtab3a

row.order<-c("good.oc","borderline.oc","poor.oc")
newtab3a.nice.names<-newtab3a.nice.names %>% arrange(match(newtab3a.nice.names$oc.levels,row.order))
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="good.oc"]<-"Good (6/18)"
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="borderline.oc"]<-"Borderline (<6/18 to 6/60)"
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="poor.oc"]<-"Poor (<6/60)"

write.table(newtab3a.nice.names,file=paste0(outdir,"/summary/data/t19.csv"),row.names=F,col.names=T,sep=",")

names(newtab3a.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab3a.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcome (PVA), count by eyes") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Outcome (PVA)" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```


```{r newtab3b, echo=FALSE}

newtab3b.nice.names<-newtab3b

row.order<-c("good.oc","borderline.oc","poor.oc")
newtab3b.nice.names<-newtab3b.nice.names %>% arrange(match(newtab3b.nice.names$oc.levels,row.order))
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="good.oc"]<-"Good (6/18)"
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="borderline.oc"]<-"Borderline (<6/18 to 6/60)"
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="poor.oc"]<-"Poor (<6/60)"

write.table(newtab3b.nice.names,file=paste0(outdir,"/summary/data/t20.csv"),row.names=F,col.names=T,sep=",")

names(newtab3b.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab3b.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcome (PinVA), count by eyes") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("Outcome (PinVA)" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("Participants not tested at 6/12",notation="symbol")

```

## Table 21, table 22

```{r outc6, echo=FALSE}

source("OUTC_6.R")

outc6a.nice.names<-male[,c("oc","male_surgery_place_gov_hospital_n","male_surgery_place_gov_hospital_percent","male_surgery_place_voluntary_hospital_n","male_surgery_place_voluntary_hospital_percent","male_surgery_place_private_hospital_n","male_surgery_place_private_hospital_percent","male_surgery_place_camp_improvised_n","male_surgery_place_camp_improvised_percent","male_surgery_place_traditional_n","male_surgery_place_traditional_percent")]

row.order<-c("good.oc","borderline.oc","poor.oc")
outc6a.nice.names<-outc6a.nice.names %>% arrange(match(outc6a.nice.names$oc,row.order))
outc6a.nice.names$oc[outc6a.nice.names$oc=="good.oc"]<-"Good (6/18)"
outc6a.nice.names$oc[outc6a.nice.names$oc=="borderline.oc"]<-"Borderline (<6/18 to 6/60)"
outc6a.nice.names$oc[outc6a.nice.names$oc=="poor.oc"]<-"Poor (<6/60)"

write.table(outc6a.nice.names,file=paste0(outdir,"/summary/data/t21.csv"),row.names=F,col.names=T,sep=",")

names(outc6a.nice.names)<-c("","n","%","n","%","n","%","n","%","n","%")

knitr::kable(outc6a.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcomes (PVA) in eyes by place of surgery (male)") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Post-surgical VA" = 1, "Gov. Hosp." = 2, "Vol. Hosp." = 2, "Priv. Hosp." = 2, "Camp Improv." = 2, "Trad." = 2),bold=T) %>%
  add_footnote(c("Gov. Hosp.: Government hospital","Vol. Hosp.: NGO hospital","Priv. Hosp.: Private hospital", "Camp Improv.: Improvised surgical camp", "Trad.: Traditional setting"), notation="symbol")

```


```{r outc6b, echo=FALSE}

outc6b.nice.names<-female[,c("oc","female_surgery_place_gov_hospital_n","female_surgery_place_gov_hospital_percent","female_surgery_place_voluntary_hospital_n","female_surgery_place_voluntary_hospital_percent","female_surgery_place_private_hospital_n","female_surgery_place_private_hospital_percent","female_surgery_place_camp_improvised_n","female_surgery_place_camp_improvised_percent","female_surgery_place_traditional_n","female_surgery_place_traditional_percent")]

row.order<-c("good.oc","borderline.oc","poor.oc")
outc6b.nice.names<-outc6b.nice.names %>% arrange(match(outc6b.nice.names$oc,row.order))
outc6b.nice.names$oc[outc6b.nice.names$oc=="good.oc"]<-"Good (6/18)"
outc6b.nice.names$oc[outc6b.nice.names$oc=="borderline.oc"]<-"Borderline (<6/18 to 6/60)"
outc6b.nice.names$oc[outc6b.nice.names$oc=="poor.oc"]<-"Poor (<6/60)"

write.table(outc6b.nice.names,file=paste0(outdir,"/summary/data/t22.csv"),row.names=F,col.names=T,sep=",")

names(outc6b.nice.names)<-c("","n","%","n","%","n","%","n","%","n","%")

knitr::kable(outc6b.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcomes (PVA) in eyes by place of surgery (female)") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Post-surgical VA" = 1, "Gov. Hosp." = 2, "Vol. Hosp." = 2, "Priv. Hosp." = 2, "Camp Improv." = 2, "Trad." = 2),bold=T) %>%
  add_footnote(c("Gov. Hosp.: Government hospital","Vol. Hosp.: NGO hospital","Priv. Hosp.: Private hospital", "Camp Improv.: Improvised surgical camp", "Trad.: Traditional setting"), notation="symbol")

```

Variation in outcome by place of surgery allows for monitoring of quality across providers. Where providers are outliers, in terms of poor quality, steps to address this should be incorporated in eye health planning.

\newpage

# Glossary of definitions

VI = Vision Impairment
PVA = Presenting Visual Acuity
PinVA = Pinhole Visual Acuity

LogMAR 0.3 = Snellen 6/12
LogMAR 0.47 = Snellen 6/18
LogMAR 1.0 = Snellen 6/60
LogMAR 1.8 = Snellen 3/60

## Bilateral VI

Blindness: PVA of logMAR >=1.8 in the better eye

Severe VI: PVA of logMAR >=1.3 in the better eye

Moderate VI: PVA of logMAR >=1.0 in the better eye

Mild VI: Participants not measured to logMAR 0.3

## Unilateral VI

Blindness: PVA of logMAR >=1.8 in one eye and logMAR 0.47 or better in the other eye

Severe VI: PVA of logMAR >=1.3 in one eye and logMAR 0.47 or better in the other eye

Moderate VI: PVA of logMAR >=1.0 in one eye and logMAR 047 or better in the other eye

Mild VI: Participants not measured to logMAR 0.3

## Cataract-related VI and surgical outcomes

Operable cataract: PinVA at <3/60, <6/60 and <6/18 thresholds plus opacity in the lens plus untreated cataract as principal cause 

Operated cataract: Aphakia or pseudoaphakia (with or without posterior capsule opacification [PCO]) 


```{r big_boi_maker, include=FALSE}

source("SUMMARY_OUT.R")

write.table(bigboi,file=paste0(outdir,"/summary/web_summary.csv"),row.names=F,col.names=T,sep=",")

```
