---
output:
 pdf_document:
  toc: FALSE
---

```{r setup, include=FALSE}

library(rmarkdown)
library(readxl)
library(knitr)
library(tinytex)
library(kableExtra)
library(float)
library(RColorBrewer)
#install_tinytex()

library(tidyverse)
library(stringr)
library(treemap)
library(maditr)
library(here)

options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = FALSE)

unlink(here("outputs",ID),recursive = T)
dir.create(here("outputs",ID))
outdir<-ID
dir.create(here("outputs",ID,"/summary"))
dir.create(here("outputs",ID,"/summary/data"))
dir.create(here("outputs",ID,"/raw"))
dir.create(here("outputs",ID,"/raw/data"))

raab_merged<-read.csv(here("data","raabs_612.csv"),check.names=F)
raab<-raab_merged[raab_merged$raab_id==ID,]
raab$date<-as.Date(raab$date)
raab$surgery_none_reason<-paste0("first_",raab$surgery_none_reason," second_",raab$surgery_none_reason_2)

pop_merged<-read.csv(here("data","raabs_pop_612_repo.csv"),check.names=F)
popfives<-pop_merged[pop_merged$raab_id==ID,]

meta<-read.csv(here("data","meta.csv"),check.names=F)

raab_meta<-meta[meta$raab_id==ID,]
PI<-raab_meta$pi_name
TITLE<-raab_meta$survey_title

write.table(raab_meta, here("outputs", ID, "raw", "meta.csv"),row.names=F,col.names=T,sep=",",na="")
write.table(raab, here("outputs", ID, "raw", "data", "survey_data.csv"),row.names=F,col.names=T,sep=",")
write.table(popfives, here("outputs", ID, "raw", "data","population.csv"),row.names=F,col.names=T,sep=",")

source(here("RAAB6_scripts","raab_functions.R"))

```

---
title: |
 <center>Rapid Assessment of Avoidable Blindness
subtitle: Report of findings from `r TITLE`
author: Principal Investigator `r PI`
output:
 pdf_document:
  toc: FALSE
tables:
 yes
header-includes:
- \usepackage{float}
- \usepackage{tabu}
- \usepackage{graphicx}
---

```{r raab_report_setup, include=FALSE}

source(here("RAAB6_scripts","RAAB6_report_setup.R"))

```

\center

Report generated: `r format(Sys.time(), '%d %B %Y')`

\raggedright

```{=latex}
\tableofcontents
```


\newpage

# Sample representation

## Table 1

```{r sum1, echo=FALSE}

source(here("RAAB6_scripts","response-rate.R"))

sum1.nice.names<-sum1

write.table(sum1.nice.names, here("outputs", ID, "summary", "data", "t1.csv"),row.names=F,col.names=T,sep=",")

names(sum1.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(sum1.nice.names, booktabs=T, linesep = "", caption = "Eligible persons, coverage, absentees and refusals") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Exam status" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("The response rate is the percent examined",notation="symbol")

```

The response rate indicates the proportion of eligible, enumerated people who were examined. The RAAB sample size calculator includes the expected non-response rate in the sample and increases the sample size accordingly. This ensures that the sample size is powerful enough to estimate the prevalence of blindness with the desired precision.

If the response rate is lower than 80-90%, there is a concern that the conditions under review in the 10-20% who were not examined may be different to those that were examined (non-response bias). For example, non-responders in a RAAB may be younger than responders (e.g., working age vs retired) and may, on average, be less likely to be vision impaired.

If the response rate is over 95%, this might be an indication that eligible participants who were absent or refused to participate were not enumerated but rather replaced by eligible participants in the next household, which would introduce selection bias and mean that results are not representative of the population. For example, people with impaired vision may be more likely to be at home and people with good vision may be more likely to be away and unavailable. In certain settings (e.g., rural or remote) a response rate over 95% is not uncommon -- participants may be more compliant with requests to stay home on the day of data collection, or more likely to work in the environment close to their home.

It is important to review this information in relation to the tables on representativeness of the sample below to identify whether a high response rate is valid.

## Table 2, table 3

```{r asa1, echo=FALSE}

source(here("RAAB6_scripts","sample-age-gender.R"))

asa1.nice.names<-asa1[,c("age.groups.tens","female.examined.n","female.examined.pct","male.examined.n","male.examined.pct","total.examined.n","total.examined.pct")]

write.table(asa1.nice.names,here("outputs", ID, "summary", "data", "t2.csv"),row.names=F,col.names=T,sep=",")

names(asa1.nice.names)<-c("Years","n","%","n","%","n","%")

knitr::kable(asa1.nice.names, booktabs=T, linesep = "", caption = "Age and sex distribution of people examined in the sample") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Age group" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)


```


```{r asa2, echo=FALSE}

source(here("RAAB6_scripts","pop-age-gender.R"))

asa2.nice.names<-asa2

write.table(asa2.nice.names,here("outputs", ID, "summary", "data", "t3.csv"),row.names=F,col.names=T,sep=",")

names(asa2.nice.names)<-c("Years","n","%","n","%","n","%")

knitr::kable(asa2.nice.names, booktabs=T, linesep = "", caption = "Total number of people aged 50+ in survey area") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Age group" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

## Plot 1, plot 2

```{r asa1_2_plots, echo=FALSE, fig.height=3.25, fig.fullwidth=T}

fasa12<-rbind(as.numeric(asa1$female.examined.pct[1:4]),as.numeric(asa2$female.pct[1:4]))
masa12<-rbind(as.numeric(asa1$male.examined.pct[1:4]),as.numeric(asa2$male.pct[1:4]))

par(mfrow=c(1,2))

bp1<-barplot(as.matrix(fasa12),beside=T,horiz=T,yaxt='n',xaxt='n',col=c("grey","purple"),border = "white",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,16))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=c(2,5,8,11,30),labels = NA,tick=F)
text(x=par("usr")[3]-6, y=c(2,5,8,11), labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col=c("purple","grey"),pch=15,c("Survey population","Census"),bty='n',cex=0.6,title = expression(bold("Survey females")),horiz=T,title.adj=0)

bp2<-barplot(as.matrix(masa12),beside=T,horiz=T,yaxt='n',xaxt='n',col=c("grey","gold"),border = "white",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,16))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=c(2,5,8,11,30),labels = NA,tick=F)
text(x=par("usr")[3]-6, y=c(2,5,8,11), labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col=c("gold","grey"),pch=15,c("Survey population","Census"),bty='n',cex=0.6,title=expression(bold("Survey males")),horiz=T,title.adj=0)

```

```{r asa1_extra_plot, echo=FALSE, fig.height=3.25, fig.fullwidth=T}

fasa1x<-as.matrix(t(asa1[1:4,c("female.examined.pct","female.nonresponse.pct")]))
masa1x<-as.matrix(t(asa1[1:4,c("male.examined.pct","male.nonresponse.pct")]))

par(mfrow=c(1,2))

bp3<-barplot(fasa1x,horiz=T,yaxt='n',xaxt='n',col=c("white","purple"),border = "purple",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,6))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=bp3[1:4],labels = NA,tick=F)
text(x=par("usr")[3]-6, y=bp3[1:4], labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col="purple",pch=c(0,15),c("Examined","Non-responders"),bty='n',cex=0.6,title = expression(bold("Enumerated females")),horiz=T,title.adj=0)

bp4<-barplot(masa1x,horiz=T,yaxt='n',xaxt='n',col=c("white","gold"),border = "gold",xlab="Proportion (%)",ylab="Age group (years)",cex.lab=0.6,ylim=c(0,6))
axis(1,at=c(-10,0,10,20,30,40,50,60,70,80,90,100),labels=c(-10,0,10,20,30,40,50,60,70,80,90,100),cex.axis=0.6)
axis(2,at=bp4[1:4],labels = NA,tick=F)
text(x=par("usr")[3]-6, y=bp4[1:4], labels=c("50-59","60-69","70-79","80+"), xpd=NA, srt=325, cex=0.6)
legend("top",col="gold",pch=c(0,15),c("Examined","Non-responders"),bty='n',cex=0.6,title=expression(bold("Enumerated males")),horiz=T,title.adj=0)

```

For your results to be useful for planning, your sample needs to be representative of the population 50 years and older. After completing the survey, we can assess representativeness by comparing the age-sex composition of the sample to the age-sex composition of the population 50 years and older.

We can also use the age-sex composition of the population 50 years and older to weight (post-stratify) crude estimates and provide age-sex adjusted (ASA) estimates for the population. We apply an 'inflation factor' -- derived from sample vs population comparisons -- to the counts of our conditions of interest in each sample age-sex group to generate extrapolated values in the population.

Often, there are more older females than younger males in the sample, and less younger males in the sample than in the population, because men are more likely to be away at work when the survey teams visit. If this is the case, use the age-sex adjusted estimates.

Important: if your sample differs from the population because one group was more likely to be unavailable (e.g., younger men, or other seasonal labourers) then you would expect to see this reflected in the response rate (i.e., more people enumerated but unavailable). If the difference between the sample and the population is high, but the proportion of people unavailable is low (e.g., response rate is still above 95%) this might be an indication that eligible participants who were absent or refused were replaced by others, which introduces bias and may mean that results are not accurate.

\newpage

# Prevalence and causes of distance vision impairment

We report the prevalence of distance vision impairment (VI) in the population 50 years and older using presenting visual acuity (PVA) in the better eye. PVA is visual acuity measured with correction, if available.

Distance vision impairment categories are defined according to the VA thresholds used in the World Health Organization's International Classification of Diseases (ICD-11).

Blindness: PVA less than 3/60 in the better eye

Severe VI: PVA less than 6/60 to 3/60 in the better eye

Moderate VI: PVA less than 6/18 to 6/60 in the better eye

Mild VI: PVA less than 6/12 to 6/18 in the better eye

## Table 4, table 5

```{r sum3, echo=FALSE}

source(here("RAAB6_scripts","crude-adj-prev-vi_RAAB6.R"))

sum3.nice.names<-sum3

sum3.nice.names$vi.level<-as.character(sum3.nice.names$vi.level)
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="blind"]<-"Blind"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="severe.vi"]<-"Severe"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="moderate.vi"]<-"Moderate"
sum3.nice.names$vi.level[sum3.nice.names$vi.level=="mild.vi"]<-"Mild"

sum3.nice.names$female.adj.ci<-paste0(sum3.nice.names$female.adj.pct.lci," - ",sum3.nice.names$female.adj.pct.uci)
sum3.nice.names$male.adj.ci<-paste0(sum3.nice.names$male.adj.pct.lci," - ",sum3.nice.names$male.adj.pct.uci)
sum3.nice.names$total.adj.ci<-paste0(sum3.nice.names$total.adj.pct.lci," - ",sum3.nice.names$total.adj.pct.uci)
sum3.nice.names$female.ci<-paste0(sum3.nice.names$female.pct.lci," - ",sum3.nice.names$female.pct.uci)
sum3.nice.names$male.ci<-paste0(sum3.nice.names$male.pct.lci," - ",sum3.nice.names$male.pct.uci)
sum3.nice.names$total.ci<-paste0(sum3.nice.names$total.pct.lci," - ",sum3.nice.names$total.pct.uci)

sum3a.nice.names<-sum3.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(sum3a.nice.names,here("outputs", ID, "summary", "data", "t4.csv"),row.names=F,col.names=T,sep=",")

names(sum3a.nice.names)<-c("","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(sum3a.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of blindness, severe, moderate and mild vision impairment") %>%
  kable_styling(full_width=T,latex_options="HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```



```{r sum4, echo=FALSE}

sum3b.nice.names<-sum3.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(sum3b.nice.names,here("outputs", ID, "summary", "data", "t5.csv"),row.names=F,col.names=T,sep=",")

names(sum3b.nice.names)<-c("","%","95% CI","Extraplotated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude")

knitr::kable(sum3b.nice.names, booktabs=T, linesep = "", caption = "Adjusted prevalence and extrapolated magnitude of blindness, severe, moderate and mild vision impairment") %>%
  kable_styling(latex_options=c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```

These tables show the crude and adjusted prevalence of vision impairment by impairment level and gender. The sample size for RAAB is calculated to provide an acceptable level of precision for the total prevalence of blindness. The accuracy of prevalence estimates for population subgroups is lower and caution should be taken in the interpretation of these data. Table 5 shows the estimated magnitude of vision impairment in the study area by gender, calculated by multiplying the crude prevalence by the population count (e.g., census data). Throughout, the 95% confidence intervals are calculated to account for RAAB's cluster sampling design.

## Table 6, table 7

```{r asa6, echo=FALSE}

source(here("RAAB6_scripts","crude-adj-cum-prev-vi_RAAB6.R"))

asa6.nice.names<-asa6
asa6.nice.names$vi.level<-as.character(asa6.nice.names$vi.level)

asa6.nice.names$vi.level[asa6.nice.names$vi.level=="mild.cumulative"]<-"Mild or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="moderate.cumulative"]<-"Moderate or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="severe.cumulative"]<-"Severe or worse"
asa6.nice.names$vi.level[asa6.nice.names$vi.level=="blind.cumulative"]<-"Blind"

asa6.nice.names$female.adj.ci<-paste0(asa6.nice.names$female.adj.pct.lci," - ",asa6.nice.names$female.adj.pct.uci)
asa6.nice.names$male.adj.ci<-paste0(asa6.nice.names$male.adj.pct.lci," - ",asa6.nice.names$male.adj.pct.uci)
asa6.nice.names$total.adj.ci<-paste0(asa6.nice.names$total.adj.pct.lci," - ",asa6.nice.names$total.adj.pct.uci)
asa6.nice.names$female.ci<-paste0(asa6.nice.names$female.pct.lci," - ",asa6.nice.names$female.pct.uci)
asa6.nice.names$male.ci<-paste0(asa6.nice.names$male.pct.lci," - ",asa6.nice.names$male.pct.uci)
asa6.nice.names$total.ci<-paste0(asa6.nice.names$total.pct.lci," - ",asa6.nice.names$total.pct.uci)

asa6a.nice.names<-asa6.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(asa6a.nice.names,here("outputs", ID, "summary", "data", "t6.csv"),row.names=F,col.names=T,sep=",")

names(asa6a.nice.names)<-c("","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(asa6a.nice.names, booktabs=T, linesep = "", caption = "Crude cumulative prevalence of blindness (any PVA <3/60), severe (any PVA <6/60), moderate (any PVA <6/18) and mild (any PVA <6/12) vision impairment") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold = T)

```



```{r asa6b,echo=FALSE}

asa6b.nice.names<-asa6.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(asa6b.nice.names,here("outputs", ID, "summary", "data", "t7.csv"),row.names=F,col.names=T,sep=",")

names(asa6b.nice.names)<-c("","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude")

knitr::kable(asa6b.nice.names, booktabs=T, linesep = "", caption = "Adjusted cumulative prevalence of blindness (any PVA <3/60), severe (any PVA <6/60), moderate (any PVA <6/18) and mild (any PVA <6/12) vision impairment") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold = T)

```


## Table 8, table 9

```{r sum6, echo=FALSE}

source(here("RAAB6_scripts","cause-vi_RAAB6.R"))

sum6.nice.names<-sum6[,1:9]

sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"1. Uncorrected refractive error"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"2. Uncorrected aphakia"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_cataract_untreated"]<-"3. Untreated cataract"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"4. Cataract surgical complications"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"5. Trachomatous corneal opacity"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"6. Other corneal opacity"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_phthisis" ]<-"7. Phthisis"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_onchocerciasis"]<-"8. Onchocerciasis"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_glaucoma" ]<-"9. Glaucoma"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"10. Diabetic retinopathy"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"11. Age-related macular degeneration"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_posterior_segment_disease" ]<-"12. Other posterior segment disease"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_myopic_degeneration" ]<-"13. Myopic degeneration"
sum6.nice.names$principal.cause[sum6.nice.names$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"14. Other globe or CNS abnomalities"

write.table(sum6.nice.names,here("outputs", ID, "summary", "data", "t8.csv"),row.names=F,col.names=T,sep=",")

sum6.for.tree<-sum6.numeric

names(sum6.nice.names) <- c(' ', 'n', '%', 'n', '%','n', '%','n', '%')

knitr::kable(sum6.nice.names, booktabs=T, linesep = "", caption = " Principal cause of blindness, severe, moderate and mild vision impairment") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(14,hline_after=T) %>%
  column_spec(1,width='6cm') %>%
  add_header_above(c("Principal cause" = 1, "Blind" = 2, "Severe" = 2, "Moderate" = 2, "Mild" = 2),bold = T)

```

```{r sum6_agg, echo=FALSE}

sum6intag<-aggregate(sum6.numeric[,2:9],by=list(sum6.numeric$cause_group_1), FUN=sum,na.rm=T)
sum6intag[nrow(sum6intag)+1,2:9]<-colSums(sum6intag[1:3,2:9])
sum6intag$Group.1[4]<-"D. Avoidable (A + B + C)"
s6tag<-aggregate(sum6.numeric[,2:9],by=list(sum6.numeric$cause_group_2), FUN=sum,na.rm=T)

sum6ag<-rbind(sum6intag,s6tag)

write.table(sum6ag,here("outputs", ID, "summary", "data", "t9.csv"),row.names=F,col.names=T,sep=",")

names(sum6ag)<-c("","n", "%", "n", "%", "n","%","n","%")

knitr::kable(sum6ag, booktabs=T, linesep = c('', '', '', '\\addlinespace'), caption = " Principal cause of blindness, severe, moderate and mild vision impairment, by intervention category") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Category" = 1, "Blind" = 2, "Severe" = 2, "Moderate" = 2, "Mild" = 2),bold = T) %>%
  add_footnote("PHC: Primary Health Care; PEC: Primary Eye Care",notation="symbol")

```


Table 8 compares the main cause of blindness, severe vision impairment, moderate vision impairment and mild visual impairment in the person. Table 9 shows what proportion of vision impairment is attributable to treatable, preventable and posterior segment disease. From these tables the priorities for intervention can be determined. The distribution of cases of blindness in the sample are visualised below.

## Plot 3

```{r treemap_1, echop=FALSE, fig.fullwidth=T}

sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"Uncorrected refractive error"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"Uncorrected aphakia"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_cataract_untreated"]<-"Untreated cataract"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"Cataract surgical complications"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"Trachomatous corneal opacity"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"Other corneal opacity"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_phthisis" ]<-"Phthisis"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_onchocerciasis"]<-"Onchocerciasis"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_glaucoma" ]<-"Glaucoma"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"Diabetic retinopathy"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"Age-related macular degeneration"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_posterior_segment_disease" ]<-"Other posterior segment disease"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_myopic_degeneration" ]<-"Myopic degeneration"
sum6.for.tree$principal.cause[sum6.for.tree$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"Other globe or CNS abnomalities"


																																																 
core_fig_cols<-brewer.pal(8,"Spectral")

treemap(sum6.for.tree[1:13,],index="principal.cause",vSize="blind.pct",type="index",palette = core_fig_cols, title="Principal causes of bilateral blindness",position.legend="right",fontsize.title=12)

```

## Table 10, table 11

```{r prev4, echo=FALSE}

source(here("RAAB6_scripts","cause-blind-gender.R"))

prev4.nice.names<-prev4[,1:7]

prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_uncorrected_refractive_error"]<-"1. Uncorrected refractive error"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_aphakia_uncorrected"]<-"2. Uncorrected aphakia"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_cataract_untreated"]<-"3. Untreated cataract"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_cataract_surgical_complications"]<-"4. Cataract surgical complications"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_trachomatous_corneal_opacity" ]<-"5. Trachomatous corneal opacity"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_corneal_opacity"]<-"6. Other corneal opacity"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_phthisis" ]<-"7. Phthisis"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_onchocerciasis"]<-"8. Onchocerciasis"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_glaucoma" ]<-"9. Glaucoma"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_diabetic_retinopathy"]<-"10. Diabetic retinopathy"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_age_related_macular_degeneration"]<-"11. Age-related macular degeneration"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_posterior_segment_disease"]<-"12. Other posterior segment disease"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_myopic_degeneration"]<-"13. Myopic degeneration"
prev4.nice.names$principal.cause[prev4.nice.names$principal.cause=="poor_vision_cause_other_globe_or_cns_abnormalities"]<-"14. Other globe or CNS abnomalities"

write.table(prev4.nice.names,here("outputs", ID, "summary", "data", "t10.csv"),row.names=F,col.names=T,sep=",")

names(prev4.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(prev4.nice.names, booktabs=T, linesep = "", caption = "Principal cause of blindness in males and females") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(14,hline_after=T) %>%
  column_spec(1,width='6cm') %>%
  add_header_above(c("Principal cause" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold = T)

```

This table shows the principal cause of blindness disaggregated by gender.

```{r prev4_agg, echo=FALSE}

prev4ag<-rbind(prev4intag,p4tag)

write.table(prev4ag,here("outputs", ID, "summary", "data", "t11.csv"),row.names=F,col.names=T,sep=",")

names(prev4ag)<-c("","n", "%", "n", "%", "n","%")

knitr::kable(prev4ag, booktabs=T, linesep = c('', '', '', '\\addlinespace'), caption = " Principal cause of blindness (PVA <3/60), by gender and intervention category") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Category" = 1, "Male" = 2, "Female" = 2, "Total" = 2),bold = T)

```


## Table 12, table 13

```{r newtab1a, echo=FALSE}

source(here("RAAB6_scripts","crude-adj-prev-vi-unilat_RAAB6.R"))

unt.nice.names<-unt

unt.nice.names$male.ci<-paste0(unt.nice.names$male.pct.lci," - ",unt.nice.names$male.pct.uci)
unt.nice.names$female.ci<-paste0(unt.nice.names$female.pct.lci," - ",unt.nice.names$female.pct.uci)
unt.nice.names$total.ci<-paste0(unt.nice.names$total.pct.lci," - ",unt.nice.names$total.pct.uci)

row.order<-c("blind.unilat","severe.unilat","moderate.unilat","mild.unilat")
unt.nice.names<-unt.nice.names %>% arrange(match(unt.nice.names$vi.level,row.order))

unt.nice.names$vi.level<-as.character(unt.nice.names$vi.level)
unt.nice.names$vi.level[unt.nice.names$vi.level=="mild.unilat"]<-"Mild in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="moderate.unilat"]<-"Moderate in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="severe.unilat"]<-"Severe in one eye only"
unt.nice.names$vi.level[unt.nice.names$vi.level=="blind.unilat"]<-"Blind in one eye only"

unta.nice.names<-unt.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]


write.table(unta.nice.names,here("outputs", ID, "summary", "data", "t12.csv"),row.names=F,col.names=T,sep=",")

names(unta.nice.names) <- c(' ', 'n', '%', '95% CI', 'n', '%', '95% CI', 'n', '%', '95% CI')

knitr::kable(unta.nice.names,booktabs=T, linesep = "",caption="Crude prevalence of blind, severe, moderate and mild unilateral vision impairment") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("VI level" = 1, "Male" = 3, "Female" = 3, "Total" = 3),bold=T) %>%
  add_footnote("Unilateral refers to cases where the other eye has PVA of 6/12",notation="symbol")

```



```{r newtab1b, echo=FALSE}


unt.nice.names$male.adj.ci<-paste0(unt.nice.names$male.adj.pct.lci," - ",unt.nice.names$male.adj.pct.uci)
unt.nice.names$female.adj.ci<-paste0(unt.nice.names$female.adj.pct.lci," - ",unt.nice.names$female.adj.pct.uci)
unt.nice.names$total.adj.ci<-paste0(unt.nice.names$total.adj.pct.lci," - ",unt.nice.names$total.adj.pct.uci)


untb.nice.names<-unt.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(untb.nice.names,here("outputs", ID, "summary", "data", "t13.csv"),row.names=F,col.names=T,sep=",")

names(untb.nice.names) <- c("", "%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magnitude")

knitr::kable(untb.nice.names, booktabs=T, linesep = "", caption="Adjusted prevalence and extrapolated magnitude of blind, severe, moderate and mild unilateral vision impairment") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) %>%
  add_footnote("Unilateral refers to cases where the other eye has PVA of 6/12",notation="symbol")

```

Cases of unilateral vision impairment acknowledge an additional sight loss burden in the population, not captured in the standard definition of bilateral vision impairment, but with the potential to impact on visual functioning.

\newpage

# Cataract

## Table 14

```{r catops, echo=FALSE}

source(here("RAAB6_scripts","crude_cat_ops_RAAB6.R"))

catops.nice.names<-catops

write.table(catops.nice.names,here("outputs", ID, "summary", "data", "t14.csv"),row.names=F,col.names=T,sep=",")

catops.nice.names$cat.ops[catops.nice.names$cat.ops=="bilateral.operated"]<-"Bilateral operated"
catops.nice.names$cat.ops[catops.nice.names$cat.ops=="unilateral.operated"]<-"Unilateral operated"
catops.nice.names$cat.ops[catops.nice.names$cat.ops=="total.operated"]<-"Total operated"

names(catops.nice.names)<-c(" ","n","%","n","%","n","%")

knitr::kable(catops.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of people bilaterally and unilaterally cataract operated") %>%
  kable_styling(full_width=T, latex_options = "HOLD_position",font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='3cm') %>%
  add_header_above(c("Oerated cataract" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

This table reports the prevalence of people in the sample with one or both eyes operated, irrespective of post-operative visual acuity or pinhole acuity in the unoperated eye.

## Table 15, table 16

```{r catcovdenom.a, echo=FALSE}

source(here("RAAB6_scripts","crude-adj-prev-cat-cov-denom_RAAB6.R"))

catcovdenom.a<-catcovdenom[,c("vi.level","female.n","female.pct","female.pct.lci","female.pct.uci","male.n","male.pct","male.pct.lci","male.pct.uci","total.n","total.pct","total.pct.lci","total.pct.uci")]

catcovdenom.a$female.ci<-paste0(catcovdenom.a$female.pct.lci," - ", catcovdenom.a$female.pct.uci)
catcovdenom.a$male.ci<-paste0(catcovdenom.a$male.pct.lci," - ", catcovdenom.a$male.pct.uci)
catcovdenom.a$total.ci<-paste0(catcovdenom.a$total.pct.lci," - ", catcovdenom.a$total.pct.uci)

catcovdenom.a.nice.names<-catcovdenom.a[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(catcovdenom.a.nice.names,here("outputs", ID, "summary", "data", "t15.csv"),row.names=F,col.names=T,sep=",")

names(catcovdenom.a.nice.names)<-c(" ","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(catcovdenom.a.nice.names, booktabs=T, linesep = "", caption = "Crude unmet need for cataract surgery at pinhole VA thresholds <3/60, <6/60, <6/18 and <6/12") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("Unmet need threshold" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```

```{r catcovdenom.b, echo=FALSE}

source(here("RAAB6_scripts","crude-adj-prev-cat-cov-denom_RAAB6.R"))

catcovdenom.b<-catcovdenom[,c("vi.level","female.adj.pct","female.adj.pct.lci","female.adj.pct.uci","male.adj.pct","male.adj.pct.lci","male.adj.pct.uci","total.adj.pct","total.adj.pct.lci","total.adj.pct.uci","extrapolated.female.n","extrapolated.male.n","extrapolated.total.n")]

catcovdenom.b$male.adj.ci<-paste0(catcovdenom.b$male.adj.pct.lci," - ",catcovdenom.b$male.adj.pct.uci)
catcovdenom.b$female.adj.ci<-paste0(catcovdenom.b$female.adj.pct.lci," - ",catcovdenom.b$female.adj.pct.uci)
catcovdenom.b$total.adj.ci<-paste0(catcovdenom.b$total.adj.pct.lci," - ",catcovdenom.b$total.adj.pct.uci)

catcovdenom.b.nice.names<-catcovdenom.b[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(catcovdenom.b.nice.names,here("outputs", ID, "summary", "data", "t16.csv"),row.names=F,col.names=T,sep=",")

names(catcovdenom.b.nice.names)<-c(" ","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude")

knitr::kable(catcovdenom.b.nice.names, booktabs=T, linesep = "", caption = "Extrapolated magnitude of unmet need for cataract surgery at pinhole VA thresholds <3/60, <6/60, <6/18 and <6/12") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("Unmet need threshold" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```

The number of people at each PinVA threshold in these tables correspond to the unmet need for cataract surgery in the calculation for effective cataract surgical coverage (see notes). People with unmet need for surgery may have cataract in one or both eyes. Those with cataract in one eye and different cause of VI in the other eye would potentially benefit from cataract surgery. Note, this group with unilateral cataract excludes people with previous cataract surgery, refractive error or uncorrected aphakia in the non-cataract eye.

## Table 17, table 18

In tables 17 & 18, the unmet need for cataract surgery is broken down by bilateral and unilateral cataract cases at each of the cataract surgical thresholds. Here, cataract cases are defined by the presence of an obvious lens opacity and assignment of cataract as the main cause of vision impairment in both eyes (bilateral) or only one eye (unilateral).

```{r asa7a, echo=FALSE}

source(here("RAAB6_scripts","adj-csc-ecsc_RAAB6.R"))
source(here("RAAB6_scripts","crude-adj-cat-vi-prev_RAAB6.R"))

asa7a.int<-asa7[,c("vi.level","female.n","female.pct","female.pct.lci","female.pct.uci","male.n","male.pct","male.pct.lci","male.pct.uci","total.n","total.pct","total.pct.lci","total.pct.uci")]

asa7a.int$female.ci<-paste0(asa7a.int$female.pct.lci," - ", asa7a.int$female.pct.uci)
asa7a.int$male.ci<-paste0(asa7a.int$male.pct.lci," - ", asa7a.int$male.pct.uci)
asa7a.int$total.ci<-paste0(asa7a.int$total.pct.lci," - ", asa7a.int$total.pct.uci)

asa7a.nice.names<-asa7a.int[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(asa7a.nice.names,here("outputs", ID, "summary", "data", "t17.csv"),row.names=F,col.names=T,sep=",")

names(asa7a.nice.names)<-c(" ","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(asa7a.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of cataract at surgical thresholds <3/60, <6/60, <6/18 and <6/12") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  pack_rows("Bilateral",1,4) %>%
  pack_rows("Unilateral",5,8) %>%
  add_header_above(c("Cataract surgical threshold" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) %>%
  add_footnote("Unilateral cases can have any level of VA in the eye without operable cataract",notation="symbol")

```


```{r asa7b, echo=FALSE}

source(here("RAAB6_scripts","adj-csc-ecsc_RAAB6.R"))
source(here("RAAB6_scripts","crude-adj-cat-vi-prev_RAAB6.R"))

asa7b.int<-asa7[,c("vi.level","female.adj.pct","female.adj.pct.lci","female.adj.pct.uci","male.adj.pct","male.adj.pct.lci","male.adj.pct.uci","total.adj.pct","total.adj.pct.lci","total.adj.pct.uci","extrapolated.female.n","extrapolated.male.n","extrapolated.total.n")]

asa7b.int$male.adj.ci<-paste0(asa7b.int$male.adj.pct.lci," - ",asa7b.int$male.adj.pct.uci)
asa7b.int$female.adj.ci<-paste0(asa7b.int$female.adj.pct.lci," - ",asa7b.int$female.adj.pct.uci)
asa7b.int$total.adj.ci<-paste0(asa7b.int$total.adj.pct.lci," - ",asa7b.int$total.adj.pct.uci)

asa7b.nice.names<-asa7b.int[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(asa7b.nice.names,here("outputs", ID, "summary", "data", "t18.csv"),row.names=F,col.names=T,sep=",")

names(asa7b.nice.names)<-c(" ","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude","Adj. %", "95% CI", "Extrapolated magnitude")

knitr::kable(asa7b.nice.names, booktabs=T, linesep = "", caption = "Adjusted prevalence and extrapolated magnitude of cataract at surgical thresholds <3/60, <6/60, <6/18 and <6/12") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  pack_rows("Bilateral",1,4) %>%
  pack_rows("Unilateral",5,8) %>%
  add_header_above(c("Cataract surgical threshold" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) %>%
  add_footnote("Unilateral cases can have any level of VA in the eye without operable cataract",notation="symbol")

```

These tables enable planning the number of surgeries required to eliminate vision impairment from cataract at a particular PinVA threshold. Assuming services will aim to operate on all eyes of people with vision impairing cataract, two surgeries are required for ‘bilateral cases’ and one surgery is required for ‘unilateral cases’.

## Table 19

```{r prev14, echo=FALSE}

source(here("RAAB6_scripts","adj-csc-ecsc_RAAB6.R"))

prev14.nice.names<-prev14[prev14$num.thresh=="csc" | prev14$num.thresh=="ecsc_612",]

prev14.nice.names$male.adjusted.ci<-paste0(prev14.nice.names$male.adjusted.lci," - ",prev14.nice.names$male.adjusted.uci)
prev14.nice.names$female.adjusted.ci<-paste0(prev14.nice.names$female.adjusted.lci," - ",prev14.nice.names$female.adjusted.uci)
prev14.nice.names$total.adjusted.ci<-paste0(prev14.nice.names$total.adjusted.lci," - ",prev14.nice.names$total.adjusted.uci)

prev14.nice.names$num.thresh<-recode_factor(prev14.nice.names$num.thresh,'csc' = "CSC",'ecsc_612' = "eCSC")

prev14a.nice.names<-prev14.nice.names[,c("num.thresh","female.adjusted","female.adjusted.ci","male.adjusted","male.adjusted.ci","total.adjusted","total.adjusted.ci","quality_gap")]

write.table(prev14a.nice.names,here("outputs", ID, "summary", "data", "t19.csv"),row.names=F,col.names=T,sep=",")

names(prev14a.nice.names)<-c("","Adj. %","95% CI","Adj. %","95% CI","Adj. %","95% CI","%")

knitr::kable(prev14a.nice.names, booktabs=T, linesep = "", caption = "Adjusted cataract surgical coverage and effective cataract surgical coverage at the person level",row.names=F) %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='4cm') %>%
  pack_rows("Cataract surgical threshold <6/12",1,2) %>%
  pack_rows("Cataract surgical threshold <6/18",3,4) %>%
  pack_rows("Cataract surgical threshold <6/60",5,6) %>%
  pack_rows("Cataract surgical threshold <3/60",7,8) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2, "Total" = 2,"Relative Quality Gap" = 1),bold=T) %>%
  add_footnote("CSC: Cataract Surgical Coverage; eCSC: Effective Cataract Surgical Coverage",notation="symbol")


```

Effective cataract surgical coverage (eCSC) measures the number of people in a population who have been operated on for cataract, and had a good outcome (at least 6/12 post-operative presenting VA), as a proportion of all people operated on or still requiring surgery. Therefore, eCSC describes service access (ie, cataract surgical coverage, [CSC]) adjusted for quality. eCSC and CSC are reported at four cataract surgical thresholds. The gap between CSC and eCSC values can be considered a quality gap; the relative quality gap is calculated as (total CSC–total eCSC)/ total CSC, with lower values reflecting better quality of cataract surgical services. See notes section for more details.

## Table 20

```{r sum11, echo=FALSE}

source(here("RAAB6_scripts","cat-barriers.R"))

sum11.nice.names<-surgery.bars[, c("Barrier","female_count","female_percent","male_count","male_percent","total_count","total_percent")]

write.table(sum11.nice.names,here("outputs", ID, "summary", "data", "t20.csv"),row.names=F,col.names=T,sep=",")

names(sum11.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(sum11.nice.names,booktabs=T, linesep = "", caption="Barriers to cataract surgery among participants with bilateral cataract and PinVA <6/60") %>%
  kable_styling(full_width=T,latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(nrow(sum11.nice.names)-1,hline_after=T) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Barrier" = 1, "Female" = 2, "Male" = 2, "Total" = 2)) %>%
  add_footnote("Participants can report 1 or 2 barriers each", notation="symbol")

```

The standard RAAB survey protocol does not allow for in-depth interviews to determine why people with cataract have not yet been operated. This preliminary data on barriers to surgery should be regarded as an indication whether more detailed qualitative studies are required.

# Surgical outcomes

Surgical outcomes are reported for all operated eyes in the sample, not at the person level. RAAB gives population based data on post-operative visual outcomes, not specific to one surgeon or one hospital and with follow-up periods ranging from months to decades.

## Table 21

```{r newtab2, echo=FALSE}

source(here("RAAB6_scripts","surg-type-outcome-gender_RAAB6.R"))

newtab2.nice.names<-newtab2

newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_iol"]<-"IOL"
newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_non_iol"]<-"Non-IOL"
newtab2.nice.names$surgery.type[newtab2.nice.names$surgery.type=="surgery_type_couching"]<-"Couching"

write.table(newtab2.nice.names,here("outputs", ID, "summary", "data", "t18.csv"),row.names=F,col.names=T,sep=",")

names(newtab2.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab2.nice.names, booktabs=T, linesep = "", caption = "Type of cataract surgery performed, count by eyes") %>%
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Surgery type" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

## Table 22, table 23

```{r newtab3a, echo=FALSE}

source(here("RAAB6_scripts","surg-type-outcome-gender_RAAB6.R"))

newtab3a.nice.names<-newtab3a

row.order<-c("good.oc","borderline.oc","poor.oc")
newtab3a.nice.names<-newtab3a.nice.names %>% arrange(match(newtab3a.nice.names$oc.levels,row.order))
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="good.oc"]<-"Good (6/12)"
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="borderline.oc"]<-"Borderline (<6/12 to 6/60)"
newtab3a.nice.names$oc.levels[newtab3a.nice.names$oc.levels=="poor.oc"]<-"Poor (<6/60)"

write.table(newtab3a.nice.names,here("outputs", ID, "summary", "data", "t22.csv"),row.names=F,col.names=T,sep=",")

names(newtab3a.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab3a.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcome (PVA), count by eyes") %>%
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("Outcome (PVA)" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```


```{r newtab3b, echo=FALSE}

newtab3b.nice.names<-newtab3b

row.order<-c("good.oc","borderline.oc","poor.oc")
newtab3b.nice.names<-newtab3b.nice.names %>% arrange(match(newtab3b.nice.names$oc.levels,row.order))
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="good.oc"]<-"Good (6/12)"
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="borderline.oc"]<-"Borderline (<6/12 to 6/60)"
newtab3b.nice.names$oc.levels[newtab3b.nice.names$oc.levels=="poor.oc"]<-"Poor (<6/60)"


write.table(newtab3b.nice.names,here("outputs", ID, "summary", "data", "t23.csv"),row.names=F,col.names=T,sep=",")

names(newtab3b.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab3b.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcome (PinVA), count by eyes") %>%
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("Outcome (PinVA)" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```


## Table 24, table 25

```{r outc6, echo=FALSE}

source(here("RAAB6_scripts","surg-place-outcome_RAAB6.R"))

outc6a.nice.names<-male[,c("oc","male_surgery_place_gov_hospital_n","male_surgery_place_gov_hospital_percent","male_surgery_place_voluntary_hospital_n","male_surgery_place_voluntary_hospital_percent","male_surgery_place_private_hospital_n","male_surgery_place_private_hospital_percent","male_surgery_place_camp_improvised_n","male_surgery_place_camp_improvised_percent","male_surgery_place_traditional_n","male_surgery_place_traditional_percent")]

row.order<-c("good.oc","borderline.oc","poor.oc")
outc6a.nice.names<-outc6a.nice.names %>% arrange(match(outc6a.nice.names$oc,row.order))
outc6a.nice.names$oc[outc6a.nice.names$oc=="good.oc"]<-"Good (6/12)"
outc6a.nice.names$oc[outc6a.nice.names$oc=="borderline.oc"]<-"Borderline (<6/12 to 6/60)"
outc6a.nice.names$oc[outc6a.nice.names$oc=="poor.oc"]<-"Poor (<6/60)"

write.table(outc6a.nice.names,here("outputs", ID, "summary", "data", "t24.csv"),row.names=F,col.names=T,sep=",")

names(outc6a.nice.names)<-c("","n","%","n","%","n","%","n","%","n","%")

knitr::kable(outc6a.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcomes (PVA) in eyes by place of surgery (male)") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Post-surgical VA" = 1, "Gov. Hosp." = 2, "Vol. Hosp." = 2, "Priv. Hosp." = 2, "Camp Improv." = 2, "Trad." = 2),bold=T)

```


```{r outc6b, echo=FALSE}

outc6b.nice.names<-female[,c("oc","female_surgery_place_gov_hospital_n","female_surgery_place_gov_hospital_percent","female_surgery_place_voluntary_hospital_n","female_surgery_place_voluntary_hospital_percent","female_surgery_place_private_hospital_n","female_surgery_place_private_hospital_percent","female_surgery_place_camp_improvised_n","female_surgery_place_camp_improvised_percent","female_surgery_place_traditional_n","female_surgery_place_traditional_percent")]

row.order<-c("good.oc","borderline.oc","poor.oc")
outc6b.nice.names<-outc6b.nice.names %>% arrange(match(outc6b.nice.names$oc,row.order))
outc6b.nice.names$oc[outc6b.nice.names$oc=="good.oc"]<-"Good (6/12)"
outc6b.nice.names$oc[outc6b.nice.names$oc=="borderline.oc"]<-"Borderline (<6/12 to 6/60)"
outc6b.nice.names$oc[outc6b.nice.names$oc=="poor.oc"]<-"Poor (<6/60)"

write.table(outc6b.nice.names,here("outputs", ID, "summary", "data", "t25.csv"),row.names=F,col.names=T,sep=",")

names(outc6b.nice.names)<-c("","n","%","n","%","n","%","n","%","n","%")

knitr::kable(outc6b.nice.names, booktabs=T, linesep = "", caption = "Post-operative visual outcomes (PVA) in eyes by place of surgery (female)") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(3,hline_after=T) %>%
  add_header_above(c("Post-surgical VA" = 1, "Gov. Hosp." = 2, "Vol. Hosp." = 2, "Priv. Hosp." = 2, "Camp Improv." = 2, "Trad." = 2),bold=T)

```

Variation in outcome by place of surgery allows for monitoring of quality across providers. Where providers are outliers, in terms of poor quality, steps to address this should be incorporated in service planning.

Gov. Hosp. = Government hospital

Vol. Hosp .= NGO hospital

Priv. Hosp. = Private hospital

Camp Improv. = Improvised surgical camp

Trad. = Traditional setting

\newpage

# Refractive error

Refractive error as a cause of vision impairment is defined as better eye PVA worse than 6/12 improving to 6/12 with pinhole. Note: As these estimates are based on presenting VA, i.e., with correction if available, participants with corrected refractive error are not included, only those with under- or uncorrected refractive error.

## Table 26, table 27

```{r new_tab_4a,echo=FALSE}

source(here("RAAB6_scripts","crude-adj-re-vi-prev_RAAB6.R"))

newtab4.nice.names<-newtab4

newtab4.nice.names$vi.level<-as.character(newtab4.nice.names$vi.level)
newtab4.nice.names$vi.level[newtab4.nice.names$vi.level=="re.blind"]<-"Blind"
newtab4.nice.names$vi.level[newtab4.nice.names$vi.level=="re.severe.vi"]<-"Severe"
newtab4.nice.names$vi.level[newtab4.nice.names$vi.level=="re.moderate.vi"]<-"Moderate"
newtab4.nice.names$vi.level[newtab4.nice.names$vi.level=="re.mild.vi"]<-"Mild"

newtab4.nice.names$female.adj.ci<-paste0(newtab4.nice.names$female.adj.pct.lci," - ",newtab4.nice.names$female.adj.pct.uci)
newtab4.nice.names$male.adj.ci<-paste0(newtab4.nice.names$male.adj.pct.lci," - ",newtab4.nice.names$male.adj.pct.uci)
newtab4.nice.names$total.adj.ci<-paste0(newtab4.nice.names$total.adj.pct.lci," - ",newtab4.nice.names$total.adj.pct.uci)
newtab4.nice.names$female.ci<-paste0(newtab4.nice.names$female.pct.lci," - ",newtab4.nice.names$female.pct.uci)
newtab4.nice.names$male.ci<-paste0(newtab4.nice.names$male.pct.lci," - ",newtab4.nice.names$male.pct.uci)
newtab4.nice.names$total.ci<-paste0(newtab4.nice.names$total.pct.lci," - ",newtab4.nice.names$total.pct.uci)

newtab4a.nice.names<-newtab4.nice.names[,c("vi.level","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(newtab4a.nice.names,here("outputs", ID, "summary", "data", "t26.csv"),row.names=F,col.names=T,sep=",")

names(newtab4a.nice.names)<-c("","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(newtab4a.nice.names, booktabs=T, linesep = "", caption = "Crude prevalence of blindness (PVA <3/60), severe (PVA <6/60), moderate (PVA <6/18) and mild (PVA <6/12) vision impairment due to refractive error") %>%
  kable_styling(latex_options=c("scale_down","HOLD_position"),font_size=8) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```


```{r new_tab_4b, echo=FALSE}

newtab4b.nice.names<-newtab4.nice.names[,c("vi.level","female.adj.pct","female.adj.ci","extrapolated.female.n","male.adj.pct","male.adj.ci","extrapolated.male.n","total.adj.pct","total.adj.ci","extrapolated.total.n")]

write.table(newtab4b.nice.names,here("outputs", ID, "summary", "data", "t27.csv"),row.names=F,col.names=T,sep=",")

names(newtab4b.nice.names)<-c("","%","95% CI","Extraplotated magnitude","%","95% CI","Extrapolated magnitude","%","95% CI","Extrapolated magintude")

knitr::kable(newtab4b.nice.names, booktabs=T, linesep = "", caption = "Adjusted prevalence and extrapolated magnitude of blindness (PVA <3/60), severe (PVA <6/60), moderate (PVA <6/18) and mild (PVA <6/12) vision impairment due to refractive error") %>%
  kable_styling(latex_options=c("scale_down","HOLD_position")) %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(c(4,7,10),width='2cm') %>%
  add_header_above(c("VI level" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```



		   

																					 
## Table 28

Effective refractive error coverage (eREC) measures the number of people in a population in need of optical correction who have received correction and had a good outcome (ie, can see at least 6/12 corrected) as a proportion of all people in need of optical correction who have accessed correction or still require it. Therefore, eREC describes service access (ie, refractive error coverage, [REC]) adjusted for quality. See notes section for more details.

Note: Both indicators refer to distance refractive error

```{r new_tab_5,echo=FALSE}

source(here("RAAB6_scripts","adj-rec-erec_RAAB6.R"))

newtab5.nice.names<-newtab5

newtab5.nice.names$rec_metric[newtab5.nice.names$rec_metric=="rec"]<-"REC"
newtab5.nice.names$rec_metric[newtab5.nice.names$rec_metric=="erec"]<-"eREC"

newtab5.nice.names$female.pct.ci<-paste0(newtab5.nice.names$female.pct.lci," - ",newtab5.nice.names$female.pct.uci)
newtab5.nice.names$female.adj.pct.ci<-paste0(newtab5.nice.names$female.adj.pct.lci," - ",newtab5.nice.names$female.adj.pct.uci)
newtab5.nice.names$male.pct.ci<-paste0(newtab5.nice.names$male.pct.lci," - ",newtab5.nice.names$male.pct.uci)
newtab5.nice.names$male.adj.pct.ci<-paste0(newtab5.nice.names$male.adj.pct.lci," - ",newtab5.nice.names$male.adj.pct.uci)
newtab5.nice.names$total.pct.ci<-paste0(newtab5.nice.names$total.pct.lci," - ",newtab5.nice.names$total.pct.uci)
newtab5.nice.names$total.adj.pct.ci<-paste0(newtab5.nice.names$total.adj.pct.lci," - ",newtab5.nice.names$total.adj.pct.uci)

newtab5.nice.names<-newtab5.nice.names[,c("rec_metric","female.adj.pct","female.adj.pct.ci","male.adj.pct","male.adj.pct.ci","total.adj.pct","total.adj.pct.ci")]

write.table(newtab5.nice.names,here("outputs", ID, "summary", "data", "t28.csv"),row.names=F,col.names=T,sep=",")

names(newtab5.nice.names) <- c(" ","Adj. %", "95% CI","Adj. %", "95% CI","Adj. %", "95% CI")

knitr::kable(newtab5.nice.names, booktabs=T, linesep = "", caption = "Adjusted distance refractive error coverage and effective refractive error coverage", row.names=F) %>%
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("eREC: Effective Refractive Error Coverage; REC: Refractive Error Coverage",notation="symbol")


```

## Table 29

```{r new_tab_6,echo=FALSE}

source(here("RAAB6_scripts","spec-use_RAAB6.R"))

newtab6.nice.names<-newtab6[1:2,]

newtab6.nice.names$specs[newtab6.nice.names$specs=="near"]<-"Near vision spectacles"
newtab6.nice.names$specs[newtab6.nice.names$specs=="distance"]<-"Distance vision
spectacles"

write.table(newtab6.nice.names,here("outputs", ID, "summary", "data", "t29.csv"),row.names=F,col.names=T,sep=",")

names(newtab6.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(newtab6.nice.names, booktabs=T, linesep = "", caption = "Distance and near vision spectacle use among study participants") %>%
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T)

```

This table reports the sample (crude) prevalence of distance and near spectacle use by gender. The categories of distance and near use are not mutually exclusive, i.e., some participants will use both. Near visual acuity is not currently measured in RAAB6; however, near vision spectacle use (for presbyopia) can be used as a proxy for refractive error coverage at near if it is assumed that 100% of the population 50 years and older will require near vision (presbyopic) correction.

\newpage


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("# Diabetic retinopathy

## Table DR1

The response rate for the diabetic retinopathy (DR) module can be lower than the standard RAAB survey if participants do not consent to a random blood sugar test (see Table 2) or to a dilated eye examination (see Table 3). Those who do not consent to dilated examination may differ from those who do. For example, they may be younger (e.g., working age rather than retired) or drivers, with the appropriate level of vision required to hold a driving licence.")

```


```{r dr-response-rate-a,echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-response-rate.R"))

dr.response.a.nice.names<-dr.response.a

write.table(dr.response.a.nice.names,here("outputs", ID, "summary", "data", "dr1.csv"),row.names=F,col.names=T,sep=",")

dr.response.a.nice.names$Exam.Status[dr.response.a.nice.names$Exam.Status=="Diabetes status assessed"]<-"Diabetes status assessed*"

names(dr.response.a.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(dr.response.a.nice.names, booktabs=T, linesep = "", caption = "RAAB survey and DR module response rate") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Exam status" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("Self-reported diabetes or consented to random blood sugar test",notation = "symbol")

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("## Table DR2")

```


```{r dr-response-rate-b,echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-response-rate.R"))

dr.response.b.nice.names<-dr.response.b

write.table(dr.response.b.nice.names,here("outputs", ID, "summary", "data", "dr2.csv"),row.names=F,col.names=T,sep=",")

dr.response.b.nice.names$Exam.Status[dr.response.b.nice.names$Exam.Status=="Suspected"]<-"Suspected*"

names(dr.response.b.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(dr.response.b.nice.names, booktabs=T, linesep = "", caption = "Known or suspected diabetes among participants assessed for diabetes status") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='5cm') %>%
  add_indent(c(2,3),level_of_indent = 1) %>%
  add_header_above(c("Exam status" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("No known history of diabetes but random blood glucose 200mg/dl or higher",notation = "symbol")

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("The proportions of diabetes that is diagnosed and undiagnosed in a population varies by region. The 2021 International Diabetes Federation Atlas (https://diabetesatlas.org/) estimates that more than half of people with diabetes are undiagnosed in Africa (53.6%), Western Pacific (52.8%) and South-East Asia (51.3%).

Note: The use of random blood glucose testing without fasting may result in over-estimates of suspected diabetes in RAAB surveys.

## Table DR3")

```


```{r dr-prev-dm-age-gender, echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-prev-dm-age-gender.R"))

dm.prev.final.nice.names<-dm.prev.final

dm.prev.final.nice.names$female.ci<-paste0(dm.prev.final.nice.names$female.pct.lci," - ", dm.prev.final.nice.names$female.pct.uci)
dm.prev.final.nice.names$male.ci<-paste0(dm.prev.final.nice.names$male.pct.lci," - ", dm.prev.final.nice.names$male.pct.uci)
dm.prev.final.nice.names$total.ci<-paste0(dm.prev.final.nice.names$total.pct.lci," - ", dm.prev.final.nice.names$total.pct.uci)

dm.prev.final.nice.names<-dm.prev.final.nice.names[,c("age.groups.tens","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(dm.prev.final.nice.names,here("outputs", ID, "summary", "data", "dr3.csv"),row.names=F,col.names=T,sep=",")

names(dm.prev.final.nice.names)<-c("Years","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(dm.prev.final.nice.names, booktabs=T, linesep = "", caption = "The prevalence of known or suspected diabetes among participants assessed for diabetes status, by age group and gender") %>% 
  kable_styling(latex_options = c("scale_down", "HOLD_position"), font_size = 8) %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  add_header_above(c("Age group" = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T)

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("The crude prevalence of diabetes among DR module participants by 10-year age group and gender is shown in Table DR2. A diagnosis of diabetes is based on either a self-reported history of diabetes (“known diabetes”), or, where diabetes is not self-reported, on a random blood sugar of 200 mg/dl or higher (“suspected diabetes”).

## Table DR4")

```


```{r dr-last-eye-exam,echo=FALSE,echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-last-eye-exam.R"))

dr.last.exam.table.nice.names<-dr.last.exam.table

write.table(dr.last.exam.table.nice.names,here("outputs", ID, "summary", "data", "t33.csv"),row.names=F,col.names=T,sep=",")

dr.last.exam.table.nice.names$last.dr.exam[dr.last.exam.table.nice.names$last.dr.exam=="dr_diabetic_last_exam_none"]<-"Never"
dr.last.exam.table.nice.names$last.dr.exam[dr.last.exam.table.nice.names$last.dr.exam=="dr_diabetic_last_exam_0_12_months"]<-"Within 1 year"
dr.last.exam.table.nice.names$last.dr.exam[dr.last.exam.table.nice.names$last.dr.exam=="dr_diabetic_last_exam_13_24_months"]<-"1-2 years"
dr.last.exam.table.nice.names$last.dr.exam[dr.last.exam.table.nice.names$last.dr.exam=="dr_diabetic_last_exam_over_24_months"]<-"More than 2 years"

names(dr.last.exam.table.nice.names)<-c("","n","%","n","%","n","%")

knitr::kable(dr.last.exam.table.nice.names, booktabs=T, linesep = "", caption = "Self-reported time since last eye examination for diabetic retinopathy among known diabetics") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  row_spec(4,hline_after=T) %>%
  column_spec(1,width='5cm') %>%
  add_header_above(c("Last exam" = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) 

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("Self-reported time since last eye examination for DR among known diabetics can be considered as a proxy for the coverage of DR (screening) services. Recall bias may affect the reliability of the findings and results should be reviewed alongside facility-based DR screening records where available.

## Table DR5")

```


```{r dr-prev-dr-in-diabetics, echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-prev-dr-in-diabetics.R"))

any.dr.mac.laser.prev.nice.names<-any.dr.mac.laser.prev

any.dr.mac.laser.prev.nice.names$female.ci<-paste0(any.dr.mac.laser.prev.nice.names$female.pct.lci," - ", any.dr.mac.laser.prev.nice.names$female.pct.uci)
any.dr.mac.laser.prev.nice.names$male.ci<-paste0(any.dr.mac.laser.prev.nice.names$male.pct.lci," - ", any.dr.mac.laser.prev.nice.names$male.pct.uci)
any.dr.mac.laser.prev.nice.names$total.ci<-paste0(any.dr.mac.laser.prev.nice.names$total.pct.lci," - ", any.dr.mac.laser.prev.nice.names$total.pct.uci)

any.dr.mac.laser.prev.nice.names<-any.dr.mac.laser.prev.nice.names[,c("condition","female.n","female.pct","female.ci","male.n","male.pct","male.ci","total.n","total.pct","total.ci")]

write.table(any.dr.mac.laser.prev.nice.names,here("outputs", ID, "summary", "data", "t34.csv"),row.names=F,col.names=T,sep=",")

any.dr.mac.laser.prev.nice.names$condition[any.dr.mac.laser.prev.nice.names$condition=="retinopathy.grade"]<-"Any retinopathy"
any.dr.mac.laser.prev.nice.names$condition[any.dr.mac.laser.prev.nice.names$condition=="maculopathy.grade"]<-"Any maculopathy"
any.dr.mac.laser.prev.nice.names$condition[any.dr.mac.laser.prev.nice.names$condition=="any.dr.mac"]<-"Any retinopathy and/or maculopathy"
any.dr.mac.laser.prev.nice.names$condition[any.dr.mac.laser.prev.nice.names$condition=="any.stdr"]<-"Sight-threatening DR (R4 and/or M2)"
any.dr.mac.laser.prev.nice.names$condition[any.dr.mac.laser.prev.nice.names$condition=="any.laser"]<-"Any laser scars"

names(any.dr.mac.laser.prev.nice.names)<-c(" ","n","%","95% CI","n","%","95% CI","n","%","95% CI")

knitr::kable(any.dr.mac.laser.prev.nice.names, booktabs=T, linesep = "", caption = "The prevalence of retinopathy and maculopathy among (known and suspect) diabetics in the sample") %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position",font_size=8)) %>%
  row_spec(0,italic=TRUE) %>%
  add_header_above(c(" " = 1, "Female" = 3, "Male" = 3, "Total" = 3),bold=T) 


```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("## Table DR6)

```


```{r dr-grade-dr-in-diabetics, echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-prev-dr-in-diabetics.R"))

ret.mac.prev.nice.names<-ret.mac.prev

ret.mac.prev.nice.names<-ret.mac.prev.nice.names[,c("grade","female.n","female.pct","male.n","male.pct","total.n","total.pct")]

write.table(ret.mac.prev.nice.names,here("outputs", ID, "summary", "data", "t35.csv"),row.names=F,col.names=T,sep=",")

ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_none"]<-"None (R0)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_mild"]<-"Mild (R1)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_observable"]<-"Observable (R2)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_referable"]<-"Referable (R3)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_proliferative"]<-"Proliferative (R4)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_retinopathy_grade_not_visualised"]<-"Not visualised (R6)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_maculopathy_grade_none"]<-"None (M0)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_maculopathy_grade_observable"]<-"Observable (M1)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_maculopathy_grade_referable"]<-"Referable (M2)"
ret.mac.prev.nice.names$grade[ret.mac.prev.nice.names$grade=="dr_maculopathy_grade_not_visualised"]<-"Not visualised (M6)"

names(ret.mac.prev.nice.names)<-c(" ","n","%","n","%","n","%")

knitr::kable(ret.mac.prev.nice.names, booktabs=T, linesep = "", caption = "Grade of retinopathy and maculopathy among (known and suspect) diabetics in the sample") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='4cm') %>%
  pack_rows("Retinopathy",1,6) %>%
  pack_rows("Maculopathy",7,10) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2, "Total" = 2),bold=T) %>%
  add_footnote("Graded using the Scottish Diabetic Retinopathy Grading Scheme", notation="symbol")

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("Crude prevalence is reported among known or suspect diabetics who consented to dilated fundus examination (see Table DR5). At the person level, the grade of retinopathy and maculopathy is calculated as the higher of right and left eye grades (R0-4 and M0-2; see Table DR6).

A ‘rule of thirds’ has previously been applied to planning for diabetic retinopathy services. This estimates that approximately one third of people with diabetes have some level of retinopathy and, of those with retinopathy, approximately one third will have sight-threatening retinopathy (STDR) requiring laser or anti-VEGF treatment. A 2021 paper by To et al. emphasises there are regional differences in DR and STDR prevalence (https://doi.org/10.1016/j.ophtha.2021.04.027). 


## Table DR7")

```


```{r dr-diabetes-vi,echo=FALSE,eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

source(here("RAAB6_scripts","dr-diabetes-vi.R"))

dm.vi.table.nice.names<-dm.vi.table

write.table(dm.vi.table.nice.names,here("outputs", ID, "summary", "data", "t36.csv"),row.names=F,col.names=T,sep=",")

dm.vi.table.nice.names$vi.level[dm.vi.table.nice.names$vi.level=="blind"]<-"Blind"
dm.vi.table.nice.names$vi.level[dm.vi.table.nice.names$vi.level=="severe.vi"]<-"Severe VI"
dm.vi.table.nice.names$vi.level[dm.vi.table.nice.names$vi.level=="moderate.vi"]<-"Moderate VI"
dm.vi.table.nice.names$vi.level[dm.vi.table.nice.names$vi.level=="mild.vi"]<-"Mild VI"

dm.vi.table.nice.names$diabetic.ci<-paste0(dm.vi.table.nice.names$diabetic.pct.lci," - ", dm.vi.table.nice.names$diabetic.pct.uci)
dm.vi.table.nice.names$non.diabetic.ci<-paste0(dm.vi.table.nice.names$non.diabetic.pct.lci," - ", dm.vi.table.nice.names$non.diabetic.pct.uci)

dm.vi.table.nice.names<-dm.vi.table.nice.names[,c("vi.level","diabetic.n","diabetic.pct","diabetic.ci","non.diabetic.n","non.diabetic.pct","non.diabetic.ci")]

names(dm.vi.table.nice.names)<-c(" ","n","%","95% CI","n","%","95% CI")

knitr::kable(dm.vi.table.nice.names, booktabs=T, linesep = "", caption = "The prevalence of vision impairment among known and suspected diabetics and non-diabetics") %>% 
  kable_styling(full_width=T, latex_options = "HOLD_position") %>%
  row_spec(0,italic=TRUE) %>%
  column_spec(1,width='4cm') %>%
  add_header_above(c("VI level" = 1, "Diabetics" = 3, "Non-diabetics" = 3),bold=T) 

```


```{r, eval=sum(!is.na(DR_check_r6$dr_diabetes_blood_consent)==TRUE)>0}

asis_output("The crude prevalence of vision impairment (due to any cause) by WHO categories among DR module participants with and without diabetes.")

```


\newpage

# Notes

## Abbreviations

CI = Confidence interval

CNS = Central nervous system

CSC = Cataract surgical coverage

eCSC = Effective cataract surgical coverage

eREC = Effective refractive error coverage

NGO = Nongovernmental organisation

PinVA = Pinhole visual acuity

PVA = Presenting visual acuity

REC = Refractive error coverage

VI = Vision impairment

## Snellen to logMAR conversion

Historic RAAB survey data variables have been updated to align with RAAB7 variable names and levels. RAAB7 records visual acuity in logMAR notation (or a code number representing light perception)


					  
0.3 = Can see 6/12

0.47 = Cannot see 6/12 but can see 6/18

1.0 = Cannot see 6/18 but can see 6/60

1.3 = Cannot see 6/60 but can see 3/60

1.8 = Cannot see 3/60 but can see 1/60

3 = Light perception

4 = No light perception

## Bilateral VI

Blindness: PVA less than 3/60 in the better eye

Severe VI: PVA less than 6/60 to 3/60 in the better eye

Moderate VI: PVA less than 6/18 to 6/60 in the better eye

Mild VI: PVA less than 6/12 to 6/18 in the better eye

## Unilateral VI

Blindness: PVA less than 3/60 in one eye, PVA 6/12 in the other eye

Severe VI: PVA less than 6/60 to 3/60 in one eye, PVA 6/12 in the other eye

Moderate VI: PVA less than 6/18 to 6/60 in one eye, PVA 6/12 in the other eye

Mild VI: PVA less than 6/12 to 6/18 in one eye, PVA 6/12 in the other eye

## Cataract-related VI and surgical outcomes

Cataract surgical threshold (operable cataract): PinVA at \<3/60, \<6/60 and \<6/18 thresholds plus lens opacity plus untreated cataract as principal cause

Operated cataract: Aphakia (excluding couched eyes) or pseudoaphakia (with or without posterior capsule opacification [PCO]) or no view of lens but cataract surgical complications as cause of vision impairment

## Cataract surgical coverage (CSC) and effective cataract surgical coverage (eCSC)

eCSC and CSC are calculated at the person level, not by eyes, and calculated at various cataract surgical thresholds.

CSC is defined as (X + Y) / (X + Y + Z)

eCSC is defined as (A + B) / (X + Y + Z)

where, e.g., at the <6/12 cataract surgical threshold:

A = individuals with unilateral operated cataract attaining 6/12 or better post-operative presenting VA in the operated eye, who have BCVA <6/12 in the other eye

B = individuals with bilateral operated cataract attaining 6/12 or better post-operative presenting VA in at least one eye

X = individuals with unilateral operated cataract (regardless of visual acuity in the operated eye) and BCVA <6/12 in the other eye

Y = individuals with bilateral operated cataract (regardless of visual acuity in the operated eyes)

Z = individuals with BCVA <6/12 in both eyes with cataract as the main cause of vision impairment in one or both eyes

For more information see: McCormick I, Butcher R, Evans JR et al. Effective cataract surgical coverage in adults aged 50 years and older: estimates from population-based surveys in 55 countries. Lancet Global Health. 2022. https://doi.org/10.1016/S2214-109X(22)00419-3 

## Refractive error coverage (REC) and effective refractive error coverage (eREC)

REC is defined as (A + B) / (A + B + C)

eREC is defined as (A) / (A + B + C)

where:

A = Individuals who present with spectacles or contact lenses for distance and whose PVA is 6/12 in the better eye (Met Need)

B = Individuals who present with spectacles or contact lenses for distance and whose PVA is <6/12 in the better eye, but who improve to 6/12 on PinVA in the better eye (Undermet Need)

C = Individuals who present without spectacles or contact lenses for distance and whose PVA is <6/12 in the better eye, but who improve to 6/12 on PinVA in the better eye (Unmet Need)

Note: This is an ‘alternative’ definition of eREC, used where uncorrected VA measurements are not available. It makes an assumption that all participants wearing correction for PVA measurement cannot see 6/12 without them and may over-estimate coverage compared to the ‘gold-standard’ calculation. For more information see: Keel S, Muller A, Block S et al. Keeping an eye on eye care: monitoring progress towards effective coverage. Lancet Global Health. 2021. https://doi.org/10.1016/S2214-109X(21)00212-6

## 95% Confidence Interval

The 95% confidence intervals are based on standard errors calculated for the clustering of the sample and the variability between clusters, specifically using the formula provided by:

Bennett S, Woods T, Liyanage WM, Smith DL. A simplified general method for cluster-sample surveys of health in developing countries. World Health Stat Q. 1991;44(3):98-106.


```{r big_boi_maker, include=FALSE}

source(here("RAAB6_scripts","SUMMARY_OUT_RAAB6.R"))

write.table(bigboi,here("outputs", ID, "summary", "web_summary.csv"),row.names=F,col.names=T,sep=",",na="")

```
